---
title: "AirQuality"
author: "Richmond Yeboah"
date: "6/6/2021"
output: html_document
---

# Load in the Dataset.

Because airquality is a pre-built dataset, we can write it to our data directory to store it for later use.

```{r}
# Install required packages
library("tidyverse")

```

# Show the data

# Look at the structure of the data

```{r}
str(airquality)

```

# Calculating Summary Statistics

If you want to look at specific statistics, here are some variations on coding. Here are 2 different ways to calculate “mean.”
```{r}
mean(airquality$Temp)
```

```{r}
mean(airquality[,4])
```

# Calculate Median, Standard Deviation, and Variance

```{r}
median(airquality$Temp)
```

```{r}
sd(airquality$Wind)
```

```{r}
var(airquality$Wind)
```

# Change the Months from 5 - 9 to May through September

```{r}
airquality$Month[airquality$Month == 5]<- "May"
airquality$Month[airquality$Month == 6]<- "June"
airquality$Month[airquality$Month == 7]<- "July"
airquality$Month[airquality$Month == 8]<- "August"
airquality$Month[airquality$Month == 9]<- "September"
```

# Look at the summary statistics of the dataset, and see how Month has change to have characters instead of numbers

```{r}
str(airquality)
```
```{r}
summary(airquality)
```

# Month is a categorical variable with different levels, called factors.

Reorder the Months so they do not default to alphabetical
```{r}
airquality$Month<-factor(airquality$Month, levels=c("May", "June","July", "August", "September"))
```

# Plot 1: Create a histogram categorized by Month with qplot

Qplot stands for “Quick-Plot” (in the ggplot2 package)


```{r}
p1 <- qplot(data = airquality,Temp,fill = Month,geom = "histogram", bins = 20)
p1
```

# Plot 2: Make a histogram using ggplot

ggplot is more sophisticated than qplot, but still uses ggplot2 package

Reorder the legend so that it is not the default (alphabetical), but rather in order that months come

Outline the bars in white using the color = “white” command

```{r}
p2 <- airquality %>%
  ggplot(aes(x=Temp, fill=Month)) +
  geom_histogram(position="identity", alpha = 0.5, binwidth = 5, color = "white")+
  scale_fill_discrete(name = "Month", labels = c("May", "June","July", "August", "September"))

p2
```

# Plot 3: Create side-by-side boxplots categorized by Month

fill=Month command fills each boxplot with a different color in the aesthetics
scale_fill_discrete makes the legend on the side for discrete color values

```{r}
p3 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Months") +
  ylab("Frequency") +
  geom_boxplot() +
  scale_fill_discrete(name = "Month", labels = c("May", "June","July", "August", "September"))
p3 
```

# Plot 4: Make the same side-by-side boxplots, but in grey-scale

Use the scale_fill_grey command for the grey-scale legend, and again, use fill=Month in the aesthetics

```{r}
p4 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Temperatures") +
  ylab("Frequency") +
  geom_boxplot()+
  scale_fill_grey(name = "Month", labels = c("May", "June","July", "August", "September"))
p4
```

# Plot 5: Now make one plot on your own of any of the variables in this dataset. It may be a scatterplot, histogram, or boxplot.

```{r}
p4 <- airquality %>%
  ggplot(aes(Month, Temp, fill = Month)) + 
  ggtitle("Temperatures") +
  xlab("Months") +
  ylab("Frequency") +
  geom_point(color="darkgreen")+
  scale_fill_discrete(name = "Month", labels = c("January", "Febuary","March", "April", "May"))
p4
```

